---
title: "Analysis"
subtitle: "PSY 4960"
author: "Peter M Grund, Helen Yu, Petra Liu"
date: "4/6/2022"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

# Setup
```{r Setup}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Import data and dependencies}
library(tidyverse)
library(mediation)
library(car)
library(viridis)
```

This open-source data file can be found at <https://osf.io/te54b/>

```{r Upload data}
df <- read_csv("In-data/emerging_adult_data.csv")

```

# Clean data

```{r Add rows}
df$ID <- 1:nrow(df) + 1000
```

```{r Select variables}
df_analysis <- df %>%
  dplyr::select(ID, mindful_1:mindful_biascheck, sex:race, school_coded, IDEA_1:`IDEA-bias-dummy`, income, stress_1:stress_bias_dummy, `moa1#1_1`:`MOA-ACH_dummy`) %>%
  # select variables of interest
  filter(age <= 25)
  # remove participants over the age of 25
```

# Score measures
## IDEA - Exploration
```{r Score IDEA}
df_analysis$exploration <- df_analysis %>% 
  dplyr::select(IDEA_1, IDEA_2, IDEA_5, IDEA_6) %>%
  # select items 1, 2, 5, and 6 to create exploration index
  rowMeans() 
  
```

## Measures of Adulthood
```{r Score measures of adulthood}
df_analysis$achievement <- df_analysis %>% 
  dplyr::select(starts_with("moa1#2")|starts_with("moa2#2")) %>% 
  # select items from the MOA to create achievement index
  rowSums()

```

## Percieved Stress Scale
```{r Score Percieved Stress Scale}
df_analysis$stress_4 <- recode(df_analysis$stress_4, '5=1; 4=2; 3=3; 2=4; 1=5')
df_analysis$stress_5 <- recode(df_analysis$stress_5, '5=1; 4=2; 3=3; 2=4; 1=5')
df_analysis$stress_7 <- recode(df_analysis$stress_7, '5=1; 4=2; 3=3; 2=4; 1=5')
df_analysis$stress_8 <- recode(df_analysis$stress_8, '5=1; 4=2; 3=3; 2=4; 1=5')
# reverse code items 4, 5, 7, and 8 on the PSS

df_analysis$stress_score <- df_analysis %>% 
  dplyr::select(stress_1:stress_10) %>% 
  rowSums()
# create single index of stress
```

## Mindfulness
```{r Score mindfulness}
df_analysis$mindful_score <- df_analysis %>%
  dplyr::select(starts_with("mindful")) %>%
  # select all rows starting with mindfulness
  rowMeans()

```

# Preliminary data analysis
```{r Summarize data}
index_summary <- df_analysis %>%
  summarize(
    mean_mindful = mean(mindful_score, na.rm = TRUE),
    sd_mindful = sd(mindful_score, na.rm = TRUE),
    mean_achievement = mean(achievement, na.rm=TRUE),
    sd_achievement = sd(achievement, na.rm=TRUE),
    mean_stress = mean(stress_score, na.rm=TRUE),
    sd_stress = sd(stress_score, na.rm=TRUE)
    ) %>%
  write_csv("Out-data/index_measure_summary.csv")

demographic_summary <- df_analysis %>%
  summarize(
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    n_males = length(which(sex == 1)),
    n_females = length(which(sex == 2)),
    n_other_gender = length(which(sex == 3)),
    n_white = length(which(race == 1)),
    n_black = length(which(race == 2)),
    n_hispanic = length(which(race == 3)),
    n_asian = length(which(race == 4)),
    n_native = length(which(race == 5)),
    n_other_race = length(which(race > 5)),
    n_male = length(which(sex == 1)),
    n_female = length(which(sex == 2))
  )


```


## Correlations
```{r}
# Between achievement and stress
cor_achieve_stress <- cor.test(df_analysis$achievement, df_analysis$stress_score)
p_cor_achieve_stress <- cor_achieve_stress$p.value

# Between stress and mindfulness
cor_mindful_stress <- cor.test(df_analysis$mindful_score, df_analysis$stress_score)
p_cor_mindful_stress <- cor_mindful_stress$p.value

# Between achievement and exploration
cor_ach_explor <- cor.test(df_analysis$achievement, df_analysis$exploration)
p_cor_ach_explor <- cor_ach_explor$p.value


```


## Linear models
```{r prelim regression analyses}
# Linear model 1: Regressing achievement score onto stress (DV and IV)
lm_1 <- lm(achievement ~ stress_score, data = df_analysis)

summary(lm_1)

# Linear model 2: Regressing exploration score onto achievement (mediator and DV)
lm2 <- lm(exploration~achievement, data=df_analysis)
summary(lm2)

#omit NA values for later mediation analysis 
df_analysis<- na.omit(df_analysis)

# Linear model 3: Regressing exploration on stress (mediator and IV) 
lm3 <- lm(exploration~stress_score, data=df_analysis)
summary(lm3)

#Linear model 4: Regressing mindfulness score onto stress (exploratory analysis)
lm4 <- lm(mindful_score~stress_score, data=df_analysis)
summary(lm4)
```

# Graphics
## Graphic of participant sample by race
```{r graphics}

by_race <- demographic_summary %>% 
  dplyr::select(n_white:n_other_race) %>%
  pivot_longer(
    cols = starts_with("n_"),
    names_prefix = "n_",
    names_to = "race",
    values_to = "number"
  ) 

by_race$race <- as.factor(by_race$race)

scale_order <- c("white", "other_race", "hispanic", "black","asian", "native")

by_race_bar <- ggplot(by_race, aes(x = race, y=number)) + 
  geom_bar(stat = "identity", aes(fill = race)) +
  theme_minimal() + 
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "none") +
  labs(title="Plot of sample by racial identity",
        x ="Race", y = "Count") +
  theme(axis.text = element_text(size=10, face="bold"))  +
  theme(legend.position="none") +
  theme(legend.text=element_text(size=10))


by_race_bar

```

## Graphic of regression of stress onto achievement 
```{r}
reg_achieve_stress <- ggplot(df_analysis, aes(x = stress_score, y = achievement)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "green") +
  labs(title="Graph of regression of stress onto achievement",
        x ="Stress score", y = "Achievement")
  
reg_achieve_stress
```


## Graphic of regression of stress onto mindfulness (exploratory analysis)
```{r}
reg_mindf_stress <- ggplot(df_analysis, aes(x = stress_score, y = mindful_score))+
  geom_point() +
  stat_smooth(method = "lm", col = "green") +
  labs(title="Graph of regression of stress onto mindfulness",
        x ="Stress score", y = "Mindfulness")

reg_mindf_stress
```

## Graphic of regression of stress on exploration 
```{r}
reg_expl_stress <- ggplot(df_analysis, aes(x = stress_score, y = exploration)) +
  geom_point() +
  stat_smooth(method= "lm", col = "green") 
```
# Incomplete Analyses and Visuals 
## Mediating Variable Analysis 
A linear regression of our IV (stress score) on the DV (achievement) has already been completed. The next step is to run a linear regression of stress score on exploration (mediating variable). From there, the Mediation package will be downloaded to combine the regression models created above and estimate medation. 

## Regression Visuals
A graphic of the regression of stress on exploration will be created as well as a graphical representation of the mediated effect. 

## Mediation Analysis 
```{r}

#linear regression of IV and mediator onto DV
lm_mediation <- lm(achievement~stress_score+exploration, df_analysis)
summary(lm_mediation)

mediation_mod<- mediate(lm3, lm_mediation, treat="stress_score", mediator="exploration", boot=T)
summary(mediation_mod)

library(tidyverse)

```